---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(caRecall)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
```

# caRecall

<!-- badges: start -->
[![R-CMD-check](https://github.com/WraySmith/caRecall/workflows/R-CMD-check/badge.svg)](https://github.com/WraySmith/caRecall/actions)
[![Codecov test coverage](https://codecov.io/gh/WraySmith/caRecall/branch/main/graph/badge.svg)](https://codecov.io/gh/WraySmith/caRecall?branch=main)
<!-- badges: end -->

The `caRecall` package is an API wrapper for the Government of Canada [Vehicle Recalls Database (VRD)](https://tc.api.canada.ca/en/detail?api=VRDB) used by the Defect Investigations and Recalls Division for vehicles, tires, and child car seats. The API wrapper provides access to recall summary information searchable by make, model, and year range, as well as detailed recall information searchable by recall number.

The package focuses on querying data from the VRD API to return the following:

- Recall summary information
- Recall counts
- Detailed recall information

## Installation

The development version of the `caRecall` package can be installed from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("WraySmith/caRecall")
```

## Usage
More documentation on `caRecall` can found at: https://wraysmith.github.io/caRecall/

Note that an API key is required to run the functions in `caRecall` and query the Vehicle Recalls Database. The key can be acquired at https://tc.api.canada.ca/en/detail?api=VRDB.

A simple example of using the `recall_by_years()` function in the package to return all recalls manufactured in 2000 could be done as follows:

```{r example}
recall_summary_2000 <- recall_by_years(start_year = 2000, end_year = 2000, limit = 3000)
tibble(recall_summary_2000)
```

The data could then be used to summarize manufacturers with the highest recall counts in the year:
```{r include=FALSE}
counts_manu_2000 <- recall_summary_2000 %>%
  group_by(`Manufacturer Name`) %>%
  tally() %>%
  filter(n >= 50)

nb.cols <- 11
mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(nb.cols)

fig <- counts_manu_2000 %>% 
  ggplot() + 
  aes(x=`Manufacturer Name`, y=n, fill=`Manufacturer Name`) +
  scale_fill_manual(values = mycolors) +
  geom_col() +
  geom_text(aes(label=n), vjust=-.5, size=3) +
  labs(y="Recall Count") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line = element_line(color = "black"),
        legend.title = element_text(size = 12),
        axis.title.x=element_blank()) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,250)) +
  ggtitle("Manufactureres with Most Recalls in 2000")
```

```{r fig1, fig.width=7, fig.height=4, fig.align='center', warning=FALSE, message=FALSE, echo=FALSE}
fig
```

Additionally, detailed recall information can be queried from the API using the `recall_details()` function:

```{r}
recall_windstar <- recall_details(1997118)
tibble(t(recall_windstar))
```

## Getting Help or Reporting an Issue

To report bugs/issues/feature requests, please file an [issue](https://github.com/WraySmith/caRecall/issues).

These are very welcome!

## How to Contribute

If you would like to contribute to the package, please see our 
[CONTRIBUTING](https://github.com/WraySmith/caRecall/blob/main/CONTRIBUTING.md) guidelines.
  
Please note that this project is released with a [Contributor Code of Conduct](https://github.com/WraySmith/caRecall/blob/main/CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.
