---
title: "Introduction to caRecall"
authors: Nathan Smith, Mitch Harris, Ryan Koenig
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vrd_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(caRecall)
```

## Introduction

The `caRecall` package is an API wrapper for the Government of Canada [Vehicle Recalls Database (VRD)](https://tc.api.canada.ca/en/detail?api=VRDB) used by the Defect Investigations and Recalls Division for vehicles, tires, and child car seats. The API wrapper provides access to recall summary information searchable by make, model, and year range, as well as detailed recall information searchable by recall number.

This vignette provides an overview of the functions available in `caRecall` and provides a demonstration of using the package to extract information from the VRD. In addition to `caRecall`, this vignette makes use of the [dplyr](https://dplyr.tidyverse.org/) and [ggplot2](https://ggplot2.tidyverse.org/) packages.

```{r warning = FALSE, message=FALSE}
library(caRecall)
library(dplyr)
library(ggplot2)
```

## API Access Key

An API key is required to run the functions in `caRecall` and query the Vehicle Recalls Database. The key can be acquired at https://tc.api.canada.ca/en/detail?api=VRDB. The API key can be set in the environment using `Sys.setenv(VRD_API = 'your_API_key_here')`. Alternatively, instead of setting the API key in the environment, the key can be passed into the `caRecall` functions as an argument.


## `caRecall` Functions

The `caRecall` package focuses on querying data from the Vehicle Recalls Database and fall into the following categories:

- Recall summary information
- Recall counts
- Detailed recall information

#### Recall Summary Information

The `recall_by_make()`, `recall_by_model()`, `recall_by_years()`, and `recall_by_number()` functions all retrieve recall summary information from the VRD based on user search terms. An overview of the search terms and function arguments includes:

- **make or model**: List of make, manufacturer, or model depending on the function used to search on. Note that the search terms are not case sensitive.
- **start_year and end_year**: The year range of the search can be specified and is based on the manufactured year and not the year a recall occurred. If the argument(s) are left empty, the search will include the start and end of the data available, respectively. Note that a vehicle year is typically a year later than the manufactured year (for example, a 2009 Ford would be entered in the database as being manufactured in 2008).
- **limit**: The VRD  API default number of items to return for a query is 25. There does not a appear to be a maximum limit associated with the API but all functions in `caRecall` also default to a limit of 25.
- **partial**: The `caRecall` functions allow the user to specify if partial matches on the seach terms are used or not (for example, "Su" would return "Suzuki", and "Subaru" matches). Note that if `partial = FALSE` is used, the number of entries returned by the functions may be less than the count provided by the 'count' functions described in the following section which return counts for all partial matches (and is the default behaviour of the VRD API. 
- **api_key**: A string of the api key needed to access the database. If left empty will error and not provide requested data frame.


If you wanted to search for the first five car recalls by the Ford make in the year 1998, you could search the following.
```{r recall_by_make}
caRecall::recall_by_make("Ford", start_year = 1998, limit = 5)
```

#### recall_by_model()

This function is very similar to the `recall_by_make()` function the only difference being that it searches by model(s) instead of make(s). 

#### recall_by_number()

Once again very similar to the previous two functions but will search by specific recall number. Not as versatile as most would not have the 7-digit recall number on hand but still available for those that do.

#### recall_by_years()

If you only wanted to search by manufacturer year then this function would best suite you.  It pulls all the recalls for any make or model and compiles it to a data frame. Lets say you wanted to search for the first 10 items with a manufacturing date between 1998 and 2001 you could use the following.
```{r recall_by_years}
caRecall::recall_by_years(start_year = 1998, end_year = 2001, limit = 10)
```

#### recall_details()

This function provides the most information on the actual recall as it state the reason for the recall.  If you have found the recall number using any of the previous functions then you can enter it (or multiples) into this function to retrieve more information about it. Lets say we use the recall number from the previous example that affected the Ford Windstar with recall number 1997118.
```{r recall_details}
t(caRecall::recall_details(1997118))
```
You can see we get several extra columns such as how many units this recall affected, what safety system it impacted, and a detailed comment explaining the recall.


### Count Functions

The next three functions are used to quickly tally up the amount of results you would get back using `recall_by_make()`, `recall_by_model()`, or `recall_by_years()`. 

#### count_recall_by_make(), count_recall_by_model(), count_recall_by_years()
These functions all have the same input parameters as their counter parts, but only return the amount of recalls obtained in the query.  Mostly useful for quick tallying. If we were to specifically look for the amount of recalls that the manufacturer Goodbaby for all of its products manufactured up to 1998 we could use the following.
```{r count_recall_by_make}
caRecall::count_recall_by_make("GOODBABY", manufacturer = TRUE, end_year = 1998)
```

### Example of How to Use the Package

#### Setting up API key

You can enter your API key for this API in manually for each query or you can automate it. First we want to check if your database has the appropriate API key in it's environment by using: 
```{r, eval = FALSE}
caRecall::get_vrd_key()
```
This should return a string with your key.  If it does not you will need to either set your API key to the environment by the function:
```{r, eval = FALSE}
Sys.setenv(VRD_API = "YOUR KEY HERE")
```
Once it is set up you should be able to leave the `api_key` parameter empty in each query.

#### Plotting

Lets say we want to look in the data base for the companies that did the most recalls with the items they manufactured in the year 2000.  First we would start with querying the API for all items recalled that were manufactured in the year 2000.
```{r recall, eval = TRUE}
recall_2000 <- caRecall::recall_by_years(start_year = 2000, end_year = 2000, limit = 10000)
```
We are putting the limit sufficiently high to assume there weren't more than 10,000 recalls that year in Canada. We can count up the data for each company by grouping by the `Manufacturer name` column and counting it. Let's look at only counts about 65 to slim down our variables.
```{r data, eval = TRUE, warning = FALSE, message=FALSE}
library(dplyr)

counts_df <- count(dplyr::group_by(recall_2000, `Manufacturer Name`))
counts_df <-counts_df[counts_df$n>=64,]
```
This reduces us to 7 companies. We could then plot this using `ggplot2` in a bar plot as seen below.
```{r plot, fig.width=7, fig.height=5, fig.align='center', warning=FALSE, message=FALSE}
library(ggplot2)

plot <- ggplot(counts_df) + 
  aes(
    x=`Manufacturer Name`,
    y=n,
    fill=`Manufacturer Name`) +
  geom_col() +
  geom_text(
    aes(label=n),
    vjust=-.5
  ) +
  labs(y="Total Recalls") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line = element_line(color = "black"),
        ) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,250)) +
  ggtitle("Top Companies by Recall Count in 2000")
  
plot
```


We could also look at how many recalls vehicles of the make Ford has done compared to Nissan for vehicles manufactured over the last 10 years.
We would start by querying a list with both "Ford" and "Nissan" in it.  We would want the start year to be 2011 and once again we would want to set the limit high enough to contain all our queries. We would need to set partial to be true in this case otherwise it will try to find makes that contain "Ford" and "Nissan".
```{r}
recall_2010 <- caRecall::recall_by_make(c("Ford","Nissan"), start_year = 2011, limit = 10000, partial = TRUE)
```

We can then plot the total amount of recalls for each manufacturers year. I would recommend reading up on the the ggplot2 package for more details on how to create these plots.

```{r plot2, fig.width=7, fig.height=5, fig.align='center', warnings = FALSE, message = FALSE}
plot2 <- recall_2010 %>%
  ggplot2::ggplot() +
  aes(x=`Year`,
      color = `Make name`) + 
  geom_point(stat='count', size=4) + 
  scale_x_continuous(breaks = seq(2010,2021,1)) +
  ggtitle("Amount of Recalls per Year")

plot2
```


### Documentation for the wrapped API can be found at:

https://tc.api.canada.ca/en/detail?api=VRDB


