---
title: "Introduction to caRecall"
authors: Nathan Smith, Mitch Harris, Ryan Koenig
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vrd_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(RColorBrewer)
```

## Introduction

The `caRecall` package is an API wrapper for the Government of Canada [Vehicle Recalls Database (VRD)](https://tc.api.canada.ca/en/detail?api=VRDB) used by the Defect Investigations and Recalls Division for vehicles, tires, and child car seats. The API wrapper provides access to recall summary information searchable by make, model, and year range, as well as detailed recall information searchable by recall number.

This vignette provides an overview of the functions available in `caRecall` and provides a demonstration of using the package to extract information from the VRD. In addition to `caRecall`, this vignette makes use of the [dplyr](https://dplyr.tidyverse.org/) and [ggplot2](https://ggplot2.tidyverse.org/) packages. The `caRecall` functions return data.frames but the vignette presents most function outputs as a tibble for presentation.

```{r warning = FALSE, message=FALSE}
library(caRecall)
library(dplyr)
library(ggplot2)
```

## API Access Key

An API key is required to run the functions in `caRecall` and query the Vehicle Recalls Database. The key can be acquired at https://tc.api.canada.ca/en/detail?api=VRDB. The API key can be set in the environment using `Sys.setenv(VRD_API = 'your_API_key_here')`. Alternatively, instead of setting the API key in the environment, the key can be passed into the `caRecall` functions as an argument.


## `caRecall` Functions

The `caRecall` package focuses on querying data from the Vehicle Recalls Database and fall into the following categories:

- Recall summary information
- Recall counts
- Detailed recall information

#### Recall Summary Information

The `recall_by_make()`, `recall_by_model()`, `recall_by_years()`, and `recall_by_number()` functions all retrieve recall summary information from the VRD based on user search terms. An overview of the search terms and function arguments includes:

- **make or model**: List of make, manufacturer, or model depending on the function used to search on. Note that the search terms are not case sensitive.
- **recall_number**: List of recall numbers to search on (however, this is typically unknown unless previous searches have been conducted). 
- **start_year and end_year**: The year range of the search can be specified and is based on the manufactured year and not the year a recall occurred. If the argument(s) are left empty, the search will include the start and end of the data available, respectively. Note that a vehicle year is typically a year later than the manufactured year (for example, a 2009 Ford would be entered in the database as being manufactured in 2008). This also means that a recall date may occur prior to the year of a vehicle.
- **limit**: The VRD  API default number of items to return for a query is 25. There does not a appear to be a maximum limit associated with the API but all functions in `caRecall` also default to a limit of 25.
- **partial**: The `caRecall` functions allow the user to specify if partial matches on the search terms are used or not (for example, "Su" would return "Suzuki", and "Subaru" matches). Note that if `partial = FALSE` is used, the number of entries returned by the functions may be less than the count provided by the `count` functions described in the following section which return counts for all partial matches (and is the default behavior of the VRD API). 
- **api_key**: If an API key is not set in the environment, an access key can be passed directly into a function.

An example of using `recall_by_model()` to search for the first five recalls by 'Ford' in 1998 would be as follows:
```{r recall_by_make}
ford_1998_recalls <- recall_by_make("Ford", start_year = 1998, limit = 5)
tibble(ford_1998_recalls)
```

Or if a user wanted to search by manufactured year and return the first 10 entries between 1999 and 2001, `recall_by_years()` could be used:
```{r recall_by_years}
recalls_1998_2001 <- recall_by_years(start_year = 1999, end_year = 2001, limit = 10)
tibble(recalls_1998_2001)
```

### Recall Counts

The `count_recall_by_make()`, `count_recall_by_model()`, and `count_recall_by_years()` all return the number of recalls available in the VRD based on the search terms. It can be used to quickly get a count of recalls prior to making a query for summary information.

For example, if a user is interesting in first determining the total number of recalls by the manufacturer 'GOODBABY' for all products manufactured up to 2008, the `count_recall_by_make()` could be used.
```{r count_recall_by_make}
caRecall::count_recall_by_make("GOODBABY", manufacturer = TRUE, end_year = 1998)
```

#### Detailed Recall Information

The `recall_details()` function provides detailed recall information for a specified recall number. The information includes details of what the recall is for (for example, 'Seats and Restraints') as well as a description on the reason for the recall. This function can only use a recall number search term and therefore likely requires prior searching using the summary recall functions to determine recall numbers of interest. 

Note that each recall number provided to the function results in a separate API call. Accordingly, the rate of API calls is limited to one call per second to stay within the API limit. For example, providing a list of 60 recall numbers to the function would take approximately 1 minute to complete.

For example, if a Ford Windstar with recall number of 1997118 was of interest:
```{r recall_details}
recall_windstar <- recall_details(1997118)
tibble(t(recall_windstar))
```

## Example
In this section, several examples will be shown using `caRecall` to extract and visualize information of interest.

### Setup API key
After getting an API access key from https://tc.api.canada.ca/en/detail?api=VRDB, the key needs to either be set in the environment (as `VRD_API`) or passed as an argument into each `caRecall` function.

To check if the VRD API key is in the environment, the following `caRecall` function can be used:
```{r, eval = FALSE}
get_vrd_key()
```

If the key is in the environment as `VRD_API` it will return the key value. If not, the function will return an error that the key is not set. The key can be set in the environment using `Sys.setenv(VRD_API = "YOUR KEY HERE")`. Once it is set, you should be able to leave the `api_key` argument empty in a `caRecall` function call.

### Getting Data from the VRD
Lets say we want to look in the VRD for the companies that had the most recalls with the items they manufactured in the year 2000. Fist we could see how many items manufactured in 2000 were recalled in the database.
```{r eval = TRUE}
count_recall_by_years(start_year = 2000, end_year = 2000)
```

We could then query the database for the summary information associated with these recalls.
```{r eval = TRUE}
recall_summary_2000 <- recall_by_years(start_year = 2000, end_year = 2000, limit = 3000)
tibble(recall_summary_2000)
```

Note that we also could have also just set the `limit` argument to a high enough value that all entries in the database would likely be included in the summary information query; however, this could also result in returning a very large dataset if the number of available entries is very high.

We can count the number of recalls for each company by grouping on the `Manufacturer name` column and summarizing the count. We'll only look at manufacturers with recall counts above 50.
```{r data}
counts_manu_2000 <- recall_summary_2000 %>%
  group_by(`Manufacturer Name`) %>%
  tally() %>%
  filter(n >= 50)

counts_manu_2000
```

There are 11 manufacturers with greater than 50 recalls for products manufactured in 2000.
```{r plot, fig.width=7, fig.height=5, fig.align='center', warning=FALSE, message=FALSE}
nb.cols <- 11
mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(nb.cols)

counts_manu_2000 %>% 
  ggplot() + 
  aes(x=`Manufacturer Name`, y=n, fill=`Manufacturer Name`) +
  scale_fill_manual(values = mycolors) +
  geom_col() +
  geom_text(aes(label=n), vjust=-.5, size=3) +
  labs(y="Recall Count") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.line = element_line(color = "black"),
        legend.title = element_text(size = 12)) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,250)) +
  ggtitle("Manufactureres with Most Recalls in 2000")
```

We could also look at how many recalls two manufacturers have had compared with each other over the over the last 10 years (2011 to 2020).

We would start by querying the API for entries with the two manufacturers. This time, we'll just set the limit to be reasonably high instead of checking the count available in the database first.
```{r}
manu_comparison_list <- c("Ford", "Toyota")
recall_comparison <- recall_by_make(manu_comparison_list, start_year = 2011, end_year = 2020, limit = 10000)
tibble(recall_comparison)
```

We can then plot the total amount of recalls for each year grouped by manufacturer.

```{r plot2, fig.width=7, fig.height=5, fig.align='center', warnings = FALSE, message = FALSE}
recall_comparison %>%
  ggplot() +
  aes(x=`Year`, color = `Make name`) + 
  geom_point(stat='count', size=4) + 
  scale_x_continuous(breaks = seq(2010,2021,1)) +
  labs(title = 'Recall Count per Year',
       y = 'Annual Count', 
       color = 'Manufacturer') +
  theme(axis.title.x=element_blank()) 
```

The plot shows a much lower number of recalls in 2020 for both manufactures; however, this is likely because there has not been sufficient time for recalls to have occurred on more recent products.

